Should I Airbnb My Property?

### 2017-03-08 abb_Writeup.Rmd
review - logit models

The analyses above show that, overall, just less than 1 out of every 3 Airbnb'd properties was actually more profitable as a short-term rental then it would have been if rented long-term over the course of the year. This proportion, however, varied by location and property type, with apartments and city-core properties showing the highest likelihood of short-erm preference. While illustrative, the above do not fully control for property characteristics or owner actions -- which are a key component of short-term rental success.  To better understand which factors are related to a higher likelihood of short-term preference, we develop a set of logistic regression models below.

@@

where $StPref$ is the probability of short-term preference, $\frac{STPref}{1-STPref}$ is the log-likelihood of short-term prefernce, $PType$ is the property type (apt or house), $BedBath$ is the bedroom and bathroom combination of the property and $\epsilon$ is an error term.

@@

The results of all three models are shown in Table 4. For $PType$ apartments are the reference category,  for $BedBath$, 1 Bed / 1Bath is the reference category, for $Submrkt$ the City-Core is the reference category and for $CancelPolicy$ `Moderate' is the reference category.  As the additional variables are added, the model fit diagnostics steadily increase, suggesting that the increased complexity of the models adds to the quality of the fit.

In Model 1, we see that when controlling for structural characteristics, Houses have a significantly lower probability of being more profitable in short-term leasing than apartments. There is very little significant difference between 1 bed / 1 bath units and the remaining categories, with the exception of 2 bed / 2 bath. This could be due to the fact that 2 bed / 2 bath units may be very attractive in the long-term market as they are a useful product for two, non-family roomates to share.  Additionally, short-term tenants may not appreciate (pay) for the second bathroom as extra bathrooms are less of amenity on holiday than in day-to-day living.

Adding the submarket variables (Model 2) shows that all areas have a much lower probability of short-term prefernce when compared to the City Core(the reference category) when controlling for structural characteristics.  The difference between the Rural and City-Core is not significant, however,  The coefficients for the structural characteristics change little from Model 1 to Model 2, speaking the robustness of those estimates. The coefficient for House properties does decrease significantly in magnitude when adding the geographic submarkets.  This is likely due to the fact that houses are concentrated more heavily in City and Suburban areas than in the City-Core and some of the impact of House is being mediated by the geographic submarkets.

Finally, in Model 3 we add in host behavior or policies. Allowing a greater number of guests per bedroom results in a higher probability of short-term preference.  This variable may be operating a proxy for additional sleeping spaces which allow for a higher nightly rate, ceteris paribus, to be charged.  The greater the minimum stay, the lower the probability of short-term preference.  Longer minimum stay may equate to lower occupancy rates as one or two night travelers are exempted from the property.  Owners with longer minimum stays may be looking to minimize costs such as cleaning and administration.  Flexible cancellation policies result in a lower probability of short-term preference, while a Strict policy has a weak, but positive relationship with short-term preference.  Strict policies do help maintain revenues, but may also scare off some potential guests.  Strict cancellation may be a luxury for the best and highly desired properties on the market and, within our model, may be acting as a proxy for some unobserved measure of quality or demand while a flexible cancellation may be signalling the opposite.

>>>>

The analyses above show that, overall, just less than one out of every four short-term rental properties was actually more profitable as a short-term rental than it would have been if rented long-term over the course of the year. This proportion, however, varied by location and property type, with apartments and city-core properties showing the highest likelihood of short-term preference. While illustrative, the above do not fully control for property characteristics or owner actions -- which are a key component of short-term rental success.  To better understand which factors are related to a higher likelihood of short-term preference, we develop a set of logistic regression models below.

@@

where $StPref$ is the probability of short-term preference, $\frac{STPref}{1-STPref}$ is the log-likelihood of short-term preference, $PType$ is the property type (apartment or house), $BedBath$ is the bedroom and bathroom combination of the property and $\epsilon$ is an error term.

@@

The results of all three models are shown in Table 4. For $PType$ Apartments are the reference category,  for $BedBath$, 1 Bed / 1Bath is the reference category, for $Submrkt$ the City-Core is the reference category and for $CancelPolicy$ Moderate is the reference category.  As the additional variables are added, the model fit diagnostics steadily increase, suggesting that the increased complexity of the models adds to the quality of the fit.

In Model 1, we see that when controlling for structural characteristics, Houses have a significantly lower probability of being more profitable in short-term leasing than apartments. There is very little significant difference between 1 bed / 1 bath units and the remaining categories, with the exception of 2 bed / 2 bath. This could be due to the fact that 2 bed / 2 bath units may be very attractive in the long-term market as they are a useful product for two, non-family roomates to share. Additionally, short-term tenants may not appreciate (pay) for the second bathroom as extra bathrooms are less of amenity during a short vacation stay than in day-to-day living.

Adding the submarket variables (Model 2) shows that all areas have a much lower probability of short-term prefernce when compared to the City Core (the reference category) when controlling for structural characteristics.  The difference between the Rural and City-Core is not significant, however. The coefficients for the structural characteristics change little from Model 1 to Model 2, speaking the robustness of those estimates. The coefficient for House properties does decrease significantly in magnitude when adding the geographic submarkets.  This is likely due to the fact that houses are concentrated more heavily in City and Suburban areas than in the City-Core and some of the impact of House is being mediated by the geographic submarkets.

Finally, in Model 3 we add in host behavior or policies. Allowing a greater number of guests per bedroom results in a higher probability of short-term preference.  This variable may be operating as proxy for additional sleeping spaces which allow for a higher nightly rate, ceteris paribus, to be charged.  The greater the minimum stay, the lower the probability of short-term preference.  Longer minimum stays may equate to lower occupancy rates as one or two night travelers are exempted from the property.  Owners with longer minimum stays may be looking to minimize costs such as cleaning and administration.  Flexible cancellation policies result in a lower probability of short-term preference, while a Strict policy has a weak, but positive relationship with short-term preference.  Strict policies do help maintain revenues, but may also scare off some potential guests.  Strict cancellation may be a luxury for the best and highly desired properties on the market and, within our model, may be acting as a proxy for some unobserved measure of quality or demand while a flexible cancellation may be signalling the opposite.
